---
title: Visualising Data in R
subtitle: Common Charts and Plots using in the NHS
---

```{r install packages}
#| include: False
install.packages("tidyverse",
                 "rmarkdown",
                 "NHSRplotthedots",
                 "rmarkdown",
                 "here",
                 "jsonlite")
```

## Line Chart

### ggplot (non-interactive)

```{r fig-linechart}
#| label: Example Line Chart

library(here)
source(here("utility", "nhs_style.R"))           # load nhsplot style
library(NHSRdatasets)
library(tidyverse)

# filter dataset
df <- ae_attendances %>% 
  filter(org_code == "RQM", type == 1)

# make plot
ggplot(df, aes(x = period, y = attendances)) +
  geom_line(colour = "#005EB8", size = 1) +
  geom_hline(yintercept = 0,
             size = 1,
             colour="#333333") +                 # bold line at x-axis
  nhs_style() +                                  # apply nhsplot style
  ylim(c(0, max(df$attendances)+1000)) +         # add space above line
  scale_y_continuous(labels = scales::comma) +   # add 1000s comma format
  labs(title = "Chelsea & Westminster Hospital",
       x = "Year",
       y = "A&E Attendances")
```

## Statistical Process Control (SPC)

The [NHSR Plot the Dots](https://github.com/nhs-r-community/NHSRplotthedots) package is built by the [NHS-R community](https://nhsrcommunity.com) to provide tools for drawing statistical process control (SPC) charts. The package supports the NHS England programme ['Making Data Count'](https://www.england.nhs.uk/publication/making-data-count/). The programme encourages boards, managers, and analyst teams to present data in ways that show change over time, and drive better understanding of indicators than 'RAG' (red, amber, green) rated reports often present.

### Installation

```r
# install from CRAN
install.packages("NHSRplotthedots")
install.packages("rmarkdown")

# Or install the development version from GitHub using {remotes} package:
# install.packages("remotes")
remotes::install_github("https://github.com/nhs-r-community/NHSRplotthedots", build_vignettes = TRUE)
```

The help-files, and vignette within this package tell you more about the possible options for controlling the charts

### Example Chart

Below is an example of the type of chart the package produces. The plot uses the `ae_attendances` dataset from the `{NHSRdatasets}` package and a bit of `{dplyr}` code to select some organisations.

```{r fig-NHSRplotthedots}
#| label: Example SPC Chart

library(NHSRplotthedots)
library(NHSRdatasets)
library(tidyverse)

df <- ae_attendances %>% 
  filter(org_code == "RQM", type == 1, period < as.Date("2018-04-01"))

df %>%
  ptd_spc(value_field = breaches, date_field = period, improvement_direction = "decrease")
```

## Help Complete this page:
We are always looking for examples and best practice to complete this documentation. See [`Contribute to this site`](https://nhsengland.github.io/nhs-r-reporting/1.5-contribute.html) for details on how to add or ammend a page on this site.

### Proposed Content:
- Bar Chart
- Pyramid Chart
- Multi-line Chart
- Targets & Benchmarks
- KPIs
- Scatter Plot
- Box Plot
- Maps